---
params:
  species_name: ""
  fGuild: ""
  hGuild: ""
  fTab: null
  hTab: null
  sTab: null
  n_pres: 0
  n_abs: 0
  auc: 0
  weight_plot_pdf: ""
  imp_plot_pdf: ""
  results_pdf: ""
  resid_hist: ""
  resid_map: ""
format:
  pdf:
    engine: knitr
    geometry:
      - top=20mm
      - bottom=20mm
      - left=20mm
      - right=20mm
editor: 
  markdown: 
    wrap: 72
---

\\begin{center} \Huge \textbf{Species Distribution Model Summary Report}
\\ \vspace{0.2cm} \Large Species: \\textit{`r params$species_name`}
\\end{center}

\vspace{-0.8cm}

## Monthly Mean Model Predictions from 1993-2023

```{r, echo=FALSE, out.width="75%", fig.align='center'}
if(file.exists(params$results_pdf)) {
  knitr::include_graphics(params$results_pdf)
} else {
  cat("Plot not found")
}
```

\newpage

## Model Residuals from 1993-2023

All residuals are calculated as the difference between the observed
presence/absence and predicted probability of occurrence for monthly
data between 1993-2023.

### Residual Distribution

```{r, echo=FALSE, out.width="85%", fig.align='center'}
if(file.exists(params$resid_hist)) {
  knitr::include_graphics(params$resid_hist)
} else {
  cat("Results plot not found at:", params$resid_hist)
}
```

\newpage

### Monthly Mean Residuals from 1993-2023

```{r, echo=FALSE, out.width="75%", fig.align='center'}
if(file.exists(params$resid_map)) {
  knitr::include_graphics(params$resid_map)
} else {
  cat("Results plot not found at:", params$resid_map)
}
```

\newpage

## Variable Importance

```{r, echo=FALSE, out.width="75%", fig.align='center'}
if(file.exists(params$imp_plot_pdf)) {
  cat(paste0("\\includegraphics[width=\\textwidth]{", params$imp_plot_pdf, "}"))
} else {
  cat("Plot not found.")
}
```

## Ensemble Model Performance

AUCs for the ensemble model were calculated using a leave-one-out
approach. Models were built with fisheries and MOM6 data from 1993-2019
and data from 2020-2023 were used to evaluate the model performance
using the AUC. The closer the AUC is to 1, the better the model
predictions are. An AUC of 0.5 or less indicates that the model is
essentially guessing when it attempts to classify presence vs absence.

The AUC for the **`r params$species_name`** ensemble was
**`r params$auc`**.

\newpage

## Model Inputs

### Fisheries Data

A total of **`r params$n_pres`** presences and **`r params$n_abs`**
absences were used to build the **`r params$species_name`** model.

Presence/absence data was collected from the following datasets:

```{r}
#| echo: false
#| output: asis

# Your input character vector
items <- params$sTab

# Render LaTeX itemize block safely
if (length(items) > 0) {
  cat("\\begin{itemize}\n")
  cat(paste0("  \\item ", items, "\n"))
  cat("\\end{itemize}\n")
}
```

### MOM6 Data

**`r params$species_name`** was classified as a **`r params$fGuild`**
and a **`r params$hGuild`** species. The following dynamic environmental
variables (not including space, time, and static environmental variables
such as bathymetry, rugosity, and distance to the coast) were therefore
used in the species distribution models:

::: {layout-ncol="2"}
```{r}
#| echo: false
#| results: asis
#| tbl-cap: "Feeding Guild Covariates"
library(knitr)
library(kableExtra)

# 1. Force the parameter back into a data frame structure
# This 'un-flattens' the YAML list back into rows and columns
df_f <- as.data.frame(params$fTab)

# 2. If it's still one wide row, we manually wrap it into 2 columns
if(ncol(df_f) > 2) {
  df_f <- data.frame(
    Long_Name = unlist(df_f[1, grepl("Long", names(df_f))]),
    Short_Name = unlist(df_f[1, grepl("Short", names(df_f))])
  )
}

# 3. Print with fixed widths to stop the overlapping
knitr::kable(
  df_f, 
  format = "latex", 
  booktabs = TRUE, 
  col.names = c("Long Name", "Short Name"),
  row.names = FALSE  # <--- This removes the "1, 2, 3" column
) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "2cm") %>%
  kable_styling(latex_options = "hold_position", font_size = 8)
```

```{r}
#| echo: false
#| results: asis
#| tbl-cap: "Habitat Guild Covariates"
library(knitr)
library(kableExtra)

# 1. Force the parameter back into a data frame structure
# This 'un-flattens' the YAML list back into rows and columns
df_f <- as.data.frame(params$hTab)

# 2. If it's still one wide row, we manually wrap it into 2 columns
if(ncol(df_f) > 2) {
  df_f <- data.frame(
    Long_Name = unlist(df_f[1, grepl("Long", names(df_f))]),
    Short_Name = unlist(df_f[1, grepl("Short", names(df_f))])
  )
}

# 3. Print with fixed widths to stop the overlapping
knitr::kable(
  df_f, 
  format = "latex", 
  booktabs = TRUE, 
  col.names = c("Long Name", "Short Name"),
  row.names = FALSE  # <--- This removes the "1, 2, 3" column
) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "2cm") %>%
  kable_styling(latex_options = "hold_position", font_size = 8)
```
:::

\vspace{-0.4cm}

## Component Model Weights

````{=tex}
\begin{minipage}[t]{0.28\textwidth}
Weights of the component models were calculated based on their respective AUCs (Area under the Curve). AUCs were calculated using observed and predicted values from 2020-2023, and then scaled to 0-1 using a Gini Index.

\textit{Note that not all sdmTMB models converged for all species, so some species may have an sdmTMB weight of zero.}
\end{minipage}
\hfill
\begin{minipage}[c]{0.68\textwidth}
\centering
```{r}
#| echo: false
#| results: asis
if(file.exists(params$weight_plot_pdf)) {
  # Use sanitize = TRUE or simply ensure the path is treated as a string
  # kableExtra's include_graphics handles this well, or raw LaTeX:
  cat(paste0("\\includegraphics[width=\\textwidth]{", params$weight_plot_pdf, "}"))
} else {
  cat("Plot not found.")
}
```
\end{minipage}
````
