---
params:
  species_name: ""
  fGuild: ""
  hGuild: ""
  fTab: null
  hTab: null
  sTab: null
  n_pres: 0
  n_abs: 0
  weight_plot_pdf: ""
  imp_plot_pdf: ""
  results_pdf: ""
  resid_hist: ""
  resid_map: ""
format:
  pdf:
    engine: knitr
    geometry:
      - top=20mm
      - bottom=20mm
      - left=20mm
      - right=20mm
editor: 
  markdown: 
    wrap: 72
---

# ***Species Distribution Model Summary Report:** `r params$species_name`*

## Monthly Mean Model Predictions from 1993-2023

```{r, echo=FALSE, out.width="80%", fig.align='center'}
if(file.exists(params$results_pdf)) {
  knitr::include_graphics(params$results_pdf)
} else {
  cat("Plot not found")
}
```

\newpage

## Model Residuals from 1993-2023

All residuals are calculated as the difference between the observed
presence/absence and predicted probability of occurrence for monthly
data between 1993-2023. Positive values indicate that the species was present, but the model predicted low probability of occurrance whereas negative values indicate that the species was absent, but the model predicted high probability of occurrance. 

### Residual Distribution
The following histogram illustrates the distribution of all residuals across the timeseries. The vertical dashed line represents the mean of all residuals. 

```{r, echo=FALSE, fig.align='center'}
if(file.exists(params$resid_hist)) {
  knitr::include_graphics(params$resid_hist)
} else {
  cat("Results plot not found at:", params$resid_hist)
}
```

\newpage

### Monthly Mean Residuals from 1993-2023
The maps below illustrate the monthly mean residuals from 1993-2023. 

```{r, echo=FALSE,  out.width="85%", fig.align='center'}
if(file.exists(params$resid_map)) {
  knitr::include_graphics(params$resid_map)
} else {
  cat("Results plot not found at:", params$resid_map)
}
```

\newpage

## Variable Importance

Variable importance is extracted from each model and normalized to their
respective sums. Variable importance for the ensemble is the weighted
average of the normalized variable importance for each component model
based on their performance.

```{r, echo=FALSE,  out.width="85%", fig.align='center'}
if(file.exists(params$imp_plot_pdf)) {
  knitr::include_graphics(params$imp_plot_pdf)
} else {
  cat("Plot not found.")
}
```

\newpage

## Model Inputs

### Fisheries Data

A total of **`r params$n_pres`** presences and **`r params$n_abs`**
absences were used to build the **`r params$species_name`** model.

Presence/absence data was collected from the following datasets:

```{r}
#| echo: false
#| output: asis

# Your input character vector
items <- params$sTab

# Render LaTeX itemize block safely
if (length(items) > 0) {
  cat("\\begin{itemize}\n")
  cat(paste0("  \\item ", items, "\n"))
  cat("\\end{itemize}\n")
}
```

### MOM6 Data

**`r params$species_name`** was classified as a **`r params$fGuild`**
and a **`r params$hGuild`** species. The following dynamic environmental
variables (not including space, time, and static environmental variables
such as bathymetry, rugosity, and distance to the coast) were therefore
used in the species distribution models:

::: {layout-ncol="2"}
```{r}
#| echo: false
#| results: asis
#| tbl-cap: "Feeding Guild Covariates"
library(knitr)
library(kableExtra)

# 1. Force the parameter back into a data frame structure
# This 'un-flattens' the YAML list back into rows and columns
df_f <- as.data.frame(params$fTab)

# 2. If it's still one wide row, we manually wrap it into 2 columns
if(ncol(df_f) > 2) {
  df_f <- data.frame(
    Long_Name = unlist(df_f[1, grepl("Long", names(df_f))]),
    Short_Name = unlist(df_f[1, grepl("Short", names(df_f))])
  )
}

# 3. Print with fixed widths to stop the overlapping
knitr::kable(
  df_f, 
  format = "latex", 
  booktabs = TRUE, 
  col.names = c("Long Name", "Short Name"),
  row.names = FALSE  # <--- This removes the "1, 2, 3" column
) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "2cm") %>%
  kable_styling(latex_options = "hold_position", font_size = 8)
```

```{r}
#| echo: false
#| results: asis
#| tbl-cap: "Habitat Guild Covariates"
library(knitr)
library(kableExtra)

# 1. Force the parameter back into a data frame structure
# This 'un-flattens' the YAML list back into rows and columns
df_f <- as.data.frame(params$hTab)

# 2. If it's still one wide row, we manually wrap it into 2 columns
if(ncol(df_f) > 2) {
  df_f <- data.frame(
    Long_Name = unlist(df_f[1, grepl("Long", names(df_f))]),
    Short_Name = unlist(df_f[1, grepl("Short", names(df_f))])
  )
}

# 3. Print with fixed widths to stop the overlapping
knitr::kable(
  df_f, 
  format = "latex", 
  booktabs = TRUE, 
  col.names = c("Long Name", "Short Name"),
  row.names = FALSE  # <--- This removes the "1, 2, 3" column
) %>%
  column_spec(1, width = "5cm") %>%
  column_spec(2, width = "2cm") %>%
  kable_styling(latex_options = "hold_position", font_size = 8)
```
:::

## Component Model Weights

Weights of the component models were calculated based on their
respective AUCs (Area under the Curve). AUCs were calculated using
observed and predicted values from 2020-2023, and then scaled to 0-1
using a Gini Index. Model weights were calculated as the fraction of the total sum of rescaled-AUCs. 

```{r, fig.align='center'}
#| echo: false
#| results: asis
if(file.exists(params$weight_plot_pdf)) {
  knitr::include_graphics(params$weight_plot_pdf)
} else {
  cat("Plot not found.")
}
```
